The temporal stability in flower visitation rates is driven by the asynchrony of pollinator abundances but does not necessarily explain plant reproductive success stability.

### 2023-05-30 Manuscript/Manuscript.Rmd
final version

---
title: "The temporal stability in flower visitation rates is driven by the asynchrony of pollinator abundances but does not necessarily explain plant reproductive success stability."
author: "Estefanía Tobajas^1, Virginia Domínguez-García^2, Francisco P. Molina^2, Ignasi Bartomeus^2"
date: '2022-10-26'
output:
  pdf_document:
    latex_engine: xelatex
  word_document: default
header-includes: \usepackage{float} \floatplacement{figure}{H} \usepackage{caption}
  \renewcommand{\caption}{Figure}
bibliography: library.bib
csl: journalofecol.csl
editor_options:
  markdown:
    wrap: sentence
---

@@

@@

![ Conceptual figure showing the expected general relationship between the visitation rate of pollinators and plant reproductive success to different species richness (R1, R2, ...). As the number of pollinator visits increases (V1, V2, …), plant reproductive success is expected to increase until that above a certain threshold, the changes in visitation rates are no longer relevant to plant reproductive success. Similarly, higer species richess is expected to increase reproductive success due to complmentary efffects among pollinators. According to this, the expected relationship between pollinator biodiversity and estability pollination depends on where we are along this curve. In point A, that temporal variation in plant reproductive success depends mainly on the variation in visitation, while in point B the reproductive success varies with the variation in richness.](Figures/conceptual_plot.png){width="400"}

@@

We analyse five plant species (*Cistus crispus*, *Cistus ladanifer*, *Cistus salviifolius*, *Halimium halimifolium*, and *Lavandula pedunculata*) that were recorded in at least 5 sites [(mean = 7 ± 0.71 sites)](lavandula%20aparece%20en%205%20sitios,%20halimiun%20en%206,%20los%20citus%20estan%20entre%207%20y%209%20sitios) in a minimum of two years per site (mean = 3.25 ± 0.16 years per plant).
The stability of pollinator visitation rates was quantified as the inverse of the coefficient of variation (mean/SD) for each plant species in each plot following Tilman (1999).
This is an adequate index given that in our dataset, as in most ecological studies, the variance scales with the mean.
To examine the role of richness and asynchrony of pollinator species abundances as drivers of stability of visitation rate, we calculate pollinator richness as the total number of different pollinator species that visit each plant species in each plot.
The synchrony index for each plant species in the plot was calculated following Loreau and de Mazancourt (2008), as: $$
\begingroup\Large
\phi =\frac{\sigma_{x_T}^2}{(\sum_{i=1}^S\sigma_{x_i})^2}
\endgroup
$$ where ${\sigma_{x_T}^2}$ is the variance of the community abundance of each plot and ${\sigma_{x_i}}$ is the standard deviation of species i abundance in a community with S species.
Note that we determined abundance as the total number of individuals that were observed, whereas the visitation rate is the total number of their visits.
This synchrony measurement varies from zero, when fluctuations are perfectly asynchronous, to one, when fluctuations are perfectly synchronous [@Loreau2008].
Finally, we express asynchrony as $1-\phi$.
We also calculated the stability of fruit proportion and the seed number for each plant species in each plot following the same methods.

@@

Stability of visitation rates ranged from 0.64 to 9.24 (mean ± SE: 1.928 ± 0.232), whereas pollinator richness ranged from 2 to 20 species (8.128 ± 0.686) and the pollinator asynchrony varied from 0 to 0.998 (0.754 ± 0.032).
There was a strong positive relationship between the stability of visitation rates and asynchrony of pollinator species abundances with greater visitation rate stability when pollinator asynchrony increased [(Estimate ± standard error: 2.903 ± 1.134; *p* = 0.017) (Fig 2A)], whereas the stability of visitation rate was not influenced by the pollinator richness (-0.001 ± 0.063; *p* = 0.986) (Fig 2B).
However, we also found a positive relationship between asynchrony and pollinator richness (0.020 ± 0.005, *p* = 0.001) (Fig 2C).
We observed one data point has a high leverage on the stability analysis (*C.crispus*, Fig 2A-B), but when we remove this point, the slope is still positive (0.968 ± 0.745).

When we further analysed each plant species separately, we observed, as expected based on the previous analysis, a consisting positive relationship between the asynchrony of pollinators and stability of visitation rate for *C. crispus*, *H. halimifoliun* and *L. pedunculata*, a weak positive relationship for *C. salviifolious* and a weak negative relationship for *C. ladanifer*.
For the relationship with pollinator richness, *C. crispus*, *C. ladanifer* and *C. salviifolius* showed a negative relationship while *H. halimifolium* and *L. pedunculata* showed positive relationships, explaining the overall lack of relationship between richness and XXXXX across all five plants (Supplementary material-Fig 1).

@@

#join plots
p_cv_visit=p1+p2+ p1.2+
  plot_annotation(tag_levels = "A")& 
  theme(plot.tag.position = c(0, 1),
        plot.tag = element_text(hjust = -0.1,vjust=0.5, size=25))& 
  theme(legend.position = "bottom",legend.background = element_blank(),legend.box.background = element_rect(colour = "black"))&theme(text = element_text(size=25))
p_cv_visit+ plot_layout(guides = "collect")

@@

To analyse whether the variability of pollinator richness and visitation rate across years affected the reproductive success of plants we used all plant species sampled.
When we analysed the variable fruit proportion, we found that the interactions between year and pollinator richness, and year and visitation rate affected the fruit proportion (richness x year: $\chi 2$ = 28.976, df = 4, *p* = \<0.0001; visitation rate x year: $\chi 2$ = 40.691, df = 4, *p* = \<0.0001).
Furthermore, we observed a significant effect of the year by itself, but not so of pollinator richness or visitation rate.
Note that we observed a general saturation of the proportion of fruits developed with most species consistently showing over 80% of fruit set, however, in years like 2017, for which the hydrological year was very dry, we found that a greater richness and visitation rate increased the fruit proportion.
We also observed a positive relationship between richness and fruit proportion in 2019.
However, in 2018, a specially wet year, the relationship between richness and fruit proportion was negative (Fig 3).

>>>>

---
title: "The temporal stability in flower visitation rates is driven by the asynchrony of pollinator abundances but does not necessarily explain plant reproductive success stability."
author: "Estefanía Tobajas^1, Virginia Domínguez-García^2, Francisco P. Molina^2, Ignasi Bartomeus^2"
date: '2022-10-26'
output:
  word_document: default
  pdf_document:
    latex_engine: xelatex
header-includes: \usepackage{float} \floatplacement{figure}{H} \usepackage{caption}
  \renewcommand{\caption}{Figure}
bibliography: library.bib
csl: journalofecol.csl
editor_options:
  markdown:
    wrap: sentence
---

@@

# Keywords
asynchrony, fruit set, pollinator richness, seed set, temporal stability.

@@

![ Conceptual figure showing the expected general relationship between the visitation rate of pollinators and plant reproductive success to different species richness (R1, R2, ...). As the number of pollinator visits increases (V1, V2, …), plant reproductive success is expected to increase until that above a certain threshold, the changes in visitation rates are no longer relevant to plant reproductive success. Similarly, higer species richess is expected to increase reproductive success due to complmentary efffects among pollinators. According to this, the expected relationship between pollinator biodiversity and stability pollination depends on where we are along this curve. In point A, that temporal variation in plant reproductive success depends mainly on the variation in visitation, while in point B the reproductive success varies with the variation in richness.](Figures/conceptual_plot.png){width="400"}

@@

We analyse five plant species (*Cistus crispus*, *Cistus ladanifer*, *Cistus salviifolius*, *Halimium halimifolium*, and *Lavandula pedunculata*) that were recorded in at least 5 sites (mean = 7 ± 0.71 sites) in a minimum of two years per site (mean = 3.25 ± 0.16 years per plant).
The stability of pollinator visitation rates was quantified as the inverse of the coefficient of variation (mean/SD) for each plant species in each plot following Tilman (1999).
This is an adequate index given that in our dataset, as in most ecological studies, the variance scales with the mean.
To examine the role of richness and asynchrony of pollinator species abundances as drivers of stability of visitation rate, we calculate pollinator richness as the total number of different pollinator species that visit each plant species in each plot.
The synchrony index for each plant species in the plot was calculated following Loreau and de Mazancourt (2008), as: $$
\begingroup\Large
\phi =\frac{\sigma_{x_T}^2}{(\sum_{i=1}^S\sigma_{x_i})^2}
\endgroup
$$ where ${\sigma_{x_T}^2}$ is the variance of the community abundance of each plot and ${\sigma_{x_i}}$ is the standard deviation of species i abundance in a community with S species.
Note that we determined abundance as the total number of individuals that were observed, whereas the visitation rate is the total number of their visits.
This synchrony measurement varies from zero, when fluctuations are perfectly asynchronous, to one, when fluctuations are perfectly synchronous [@Loreau2008].
Finally, we express asynchrony as $1-\phi$.
We also calculated the stability of fruit proportion and the seed number for each plant species in each plot following the same methods.

@@

Stability of visitation rates ranged from 0.64 to 9.24 (mean ± SE: 1.928 ± 0.232), whereas pollinator richness ranged from 2 to 20 species (8.128 ± 0.686) and the pollinator asynchrony varied from 0 to 0.998 (0.754 ± 0.032).
There was a strong positive relationship between the stability of visitation rates and asynchrony of pollinator species abundances with greater visitation rate stability when pollinator asynchrony increased (Estimate ± standard error: 2.903 ± 1.134; *p* = 0.017) (Fig 2A), whereas the stability of visitation rate was not influenced by the pollinator richness (-0.001 ± 0.063; *p* = 0.986) (Fig 2B).
However, we also found a positive relationship between asynchrony and pollinator richness (0.020 ± 0.005, *p* = 0.001) (Fig 2C).
We observed one data point has a high leverage on the stability analysis (*C.crispus*, Fig 2A-B), but when we remove this point, the slope is still positive (0.968 ± 0.745).

When we further analysed each plant species separately, we observed, as expected based on the previous analysis, a consisting positive relationship between the asynchrony of pollinators and stability of visitation rate for *C. crispus*, *H. halimifoliun* and *L. pedunculata*, a weak positive relationship for *C. salviifolious* and a weak negative relationship for *C. ladanifer*.
For the relationship with pollinator richness, *C. crispus*, *C. ladanifer* and *C. salviifolius* showed a negative relationship while *H. halimifolium* and *L. pedunculata* showed positive relationships, explaining the overall lack of relationship between he richness and stability of visitation rates across all five plants (Supplementary material-Fig 1).

@@

#join plots
p_cv_visit=p1+p2+ p1.2+
  plot_annotation(tag_levels = "A")& 
  theme(plot.tag.position = c(0, 1),
        plot.tag = element_text(hjust = -0.1,vjust=0.5, size=25))& 
  theme(legend.position = "bottom",legend.background = element_blank(),legend.box.background = element_rect(colour = "black"))&theme(text = element_text(size=25))

p_cv_visit+ plot_layout(guides = "collect")

@@

To analyse whether the variability of pollinator richness and visitation rate across years affected the reproductive success of plants we used all plant species sampled.
When we analysed the variable fruit proportion, we found that the interactions between year and pollinator richness, and year and visitation rate affected the fruit proportion (richness x year: $\chi 2$ = 28.976, df = 4, *p* = \<0.0001; visitation rate x year: $\chi 2$ = 40.691, df = 4, *p* = \<0.0001).
Furthermore, we observed a significant effect of the year by itself, but not so of pollinator richness or visitation rate.
Note that we observed a general saturation of the proportion of fruits developed with most species consistently showing over 80% of fruit set, however, in years like 2017, for which the hydrological year was very dry, we found that a greater richness and visitation rate increased the fruit proportion.
We also observed a positive relationship between richness and fruit proportion in 2019.
However, in 2018, a specially wet year, the relationship between richness and fruit proportion was negative (Fig 4).
